---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# parcel

<!-- badges: start -->
[![R-CMD-check](https://github.com/geomarker-io/parcel/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/geomarker-io/parcel/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->


The goal of parcel is to provide tools for matching real-world addresses to reference sets of addresses; e.g., "352 Helen Street", "352 Helen St." or "352 helen st". This package is motivated by the included example data resource of auditor parcel tax data from Hamilton County, Ohio.

With this specific goal in mind, parcel includes:

- functions for cleaning and tagging components of addresses: **`clean_address()`**, **`tag_address()`**, and **`create_address_stub()`**
- the `cagis_parcels` CoDEC tabular-data-resource
- functions for joining addresses to parcel identifiers based on an included model pretrained on electronic health record addresses in Hamilton County, OH: **`link_parcel()`**

## Installation

{parcel} requires the `usaddress` python module for tagging addresses and creating address stubs, as well as the `dedupe` and `dedupe-variable-address` python modules for matching addresses, all of which can be installed from inside R with:

```{r}
#| eval: false
reticulate::py_install("usaddress", pip = TRUE)
reticulate::py_install("dedupe", pip = TRUE)
reticulate::py_install("dedupe-variable-address", pip = TRUE)
```

`reticulate::py_install()` assumes a non-system version of Python is already installed.
If not, it will ask you to install one via Miniconda. Don't do this; use virtualenv instead:

```{r}
#| eval: false
library(reticulate)
install_python("3.9.12")
virtualenv_create("r-parcel", version = "3.9.12")
use_virtualenv("r-parcel")
py_install("usaddress", pip = TRUE)
py_module_available("usaddress")
```

To help R find this virtualenv for every session, either `use_virtualenv("r-parcel")` should be run in every R session *or*, better yet, by setting the `RETICULATE_PYTHON` environment variable in a user- or project-specific `.Renviron` file. You can check on which python installation chosen to be used by reticulate and why by using:

```{r}
#| eval: false
reticulate::py_config()
reticulate::py_list_packages()
```

The development version of parcel can be installed with:

```{r}
#| eval: false
renv::install("geomarker-io/parcel")
```

## Example Usage

Use `get_parcel_data()` to get the corresponding parcel data for a vector of addresses:

```{r}
#| cache: true
library(parcel)
get_parcel_data(c("1069 Overlook Avenue Cincinnati OH 45238",
                  "419 Elm St. Cincinnati OH 45238",
                  "3544 Linwood Av Cincinnati OH 45226"))
```

## CAGIS Parcels Data Details

The CAGIS Parcels tabular data resource is created using the `01_make_cagis_parcels.R` script and stored in the package.  It can be loaded using {[`codec`](https://geomarker.io/codec)}:

```{r}
d_parcel <- codec::read_tdr_csv(fs::path_package("parcel", "cagis_parcels"))

head(d_parcel)

# without CODECtools:
# read.csv(fs::path_package("parcel", "cagis_parcels"))
```

```{r}
#| results: asis
options(knitr.kable.NA = '')
codec::glimpse_schema(d_parcel) |>
  knitr::kable()
```

### Inclusion/Exclusion Criteria for Parcel Data

Auditor parcel-level data were excluded if they (1) did not contain a parcel identifier, (2) did not contain a property address number/name, or (3) were complete duplicates.

Parcels with the following land use categories are included in the data resource and others are excluded.  These were selected to reflect *residential* usages of parcels.

```{r}
library(dplyr, warn.conflicts = FALSE)

d_parcel |>
  group_by(land_use) |>
  summarize(n_parcels = n()) |>
  arrange(desc(n_parcels)) |>
  knitr::kable()
```

### Estimating the number of households per parcel

Certain calculations needs to be weighted by households instead of parcel; e.g. "What fraction of families live near roadway in Avondale?".  We assume the following as a conservative estimate of the number of households per parcel for each `land_use` code:

|`land_use`                        |n households|
|:-------------------------------|------:|
|single family dwelling          |1|
|condominium unit                |1|
|two family dwelling             |2|
|three family dwelling           |3|
|apartment, 4-19 units           |4|
|apartment, 20-39 units          |20|
|apartment, 40+ units            |40|
|landominium                     |1|
|charities, hospitals, retir     |1|
|condo or pud garage             |1|
|metropolitan housing authority  |1|
|office / apartment over         |1|
|manufactured home               |1|
|other commercial housing        |1|
|nursing home / private hospital |1|
|mobile home / trailer park      |1|
|single fam dw 0-9 acr           |1|
|independent living (seniors)    |1|
|lihtc res                       |1|
|condominium office building     |0|
|other residential structure     |0|
|boataminium                     |0|
|resid unplat 10-19 acres        |0|
|resid unplat 20-29 acres        |0|
|resid unplat 30-39 acres        |0|
|residential vacant land         |0|

### Identifiers for Parcels and Properties

A `parcel_id` refers to the Hamilton County Auditor's "Parcel Number", which is referred to as the "Property Number" within the CAGIS Open Data and uniquely identifies properties. In rare cases, multple addresses can share the same parcel boundaries, but have unique `parcel_id`s and in these cases, their resulting centroid coordinates would also be identical.

Because "second line" address components (e.g., "Unit 2B") are not captured, a single address can refer to multiple parcels in the case of condos or otherwise shared building ownership. For example, the address "323 Fifth St" has six distinct `parcel_id`s, each with different home values and land uses:

|parcel_id   | market_total_value|land_use                    |
|:-----------|------------------:|:---------------------------|
|14500010321 |             397500|condominium unit            |
|14500010317 |             123000|condominium office building |
|14500010320 |             180000|condominium unit            |
|14500010319 |             255000|condominium unit            |
|14500010322 |             388230|condominium unit            |
|14500010318 |             239500|condominium unit            |
